---
title: "Upper Quartile Expression Plots"
output: html_document
date: "2024-09-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries & Import Data

```{r prep, echo=FALSE}
library(tidyverse)
library(readxl)
library(ggtext)
library(ggh4x) # facet_nested()

PopNameOrder <- c(
  "CD4_TN", "CD4_TSCM", "CD4_Tfh", "CD4_TCM_Th1_2DN", "CD4_TCM_Th1", "CD4_TCM_Th2", 
  "CD4_TCM_Th1_2DP", "CD4_TCM_Th17", "CD4_TCM_Th1_17", "CD4_TCM_Th17DN", "CD4_TCM_Th17DP", 
  "CD4_TEM_Th1_2DN", "CD4_TEM_Th1", "CD4_TEM_Th2", "CD4_TEM_Th1_2DP", "CD4_TEM_Th17", 
  "CD4_TEM_Th1_17", "CD4_TEM_Th17DN", "CD4_TEM_Th17DP", "CD4_TEMRA", "CD4_Treg", 
  "CD8_TN", "CD8_TSCM", "CD8_Tfh", "CD8_TCM_Tc1_2DN", "CD8_TCM_Tc1", "CD8_TCM_Tc2", 
  "CD8_TCM_Tc1_2DP", "CD8_TCM_Tc17", "CD8_TCM_Tc1_17", "CD8_TCM_Tc17DN", "CD8_TCM_Tc17DP", 
  "CD8_TEM_Tc1_2DN", "CD8_TEM_Tc1", "CD8_TEM_Tc2", "CD8_TEM_Tc1_2DP", "CD8_TEM_Tc17", 
  "CD8_TEM_Tc1_17", "CD8_TEM_Tc17DN", "CD8_TEM_Tc17DP", "CD8_TEMRA", "CD8_Treg")

PopLabelOrder <- PopNameOrder %>% 
  str_replace_all("_", " ") %>% 
  str_replace_all(" T.1 2DN", "") %>% 
  str_replace_all("1 2", "1/2") %>% 
  str_replace_all("1 17", "1/17")

expressiondata <- "data/expressiondata.csv" %>% 
  read_csv() %>% 
  mutate(PTID = as.factor(PTID),
       Visit = as.factor(Visit),
       Infection = as.factor(Infection),
       FollowUp = as.factor(FollowUp),
       Ascension = as.factor(Ascension),
       PopName = as.factor(PopName),
       PopIndex = as.factor(PopIndex)) %>% 
  mutate(PopName = factor(PopName, levels = PopNameOrder))

allupperqdata <- "data/allupperqdata.csv" %>%
  read_csv() %>% 
  mutate(PTID = as.factor(PTID),
     Visit = as.factor(Visit),
     Infection = as.factor(Infection),
     FollowUp = as.factor(FollowUp),
     Ascension = as.factor(Ascension),
     PopName = as.factor(PopName),
     PopIndex = as.factor(PopIndex)) %>% 
  mutate(PopName = factor(PopName, levels = PopNameOrder)) %>% 
  mutate(PopLabel = factor(PopLabel, levels = PopLabelOrder))
```

# Uninf vs CT+ vs CT+CoP

## Build plot function

```{r build function}

plotinf <- function(boxdata, jitterdata, aspect_ratio, y_variable) {
  plot <- boxdata %>%
    ggplot(aes(x = PopLabel, 
               y = !!sym(y_variable),  # Convert y_variable to symbol
               fill = Infection)) +
    geom_boxplot(color = "black", 
                 outlier.shape = NA, 
                 coef = 0, 
                 size = 0.25) +
    scale_fill_manual(values = c("#7CAE00", "#F8766D", "#CD9600"),
                      labels = c("Uninf", "CT+", "CT+CoP")) +
    geom_jitter(data = jitterdata, 
                aes(x = PopLabel, 
                    y = !!sym(y_variable),  # Convert y_variable to symbol
                    color = Infection),
                size = 0.1,
                position = position_jitterdodge(jitter.width = 0.15, 
                                                dodge.width = 0.75)) +  
    scale_color_manual(values = c("#3F5E00", "#A83C34", "#7D6000"), 
                       guide = "none") +
    coord_cartesian(ylim = c(NA, max(jitterdata[[y_variable]]))) + 
    labs(y = y_variable) +
    theme_classic() +
    theme(aspect.ratio = aspect_ratio,
          axis.title = element_text(size = 11, face = "bold"),  # Increase axis title font size
          axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text = element_text(size = 11, color = "black"),
          legend.text = element_text(size = 11),
          legend.title = element_blank(),
          legend.background = element_rect(fill = "transparent"),
          legend.box.margin = margin(0, 0, 0, 0),
          legend.spacing.x = unit(0, "mm"),
          legend.spacing.y = unit(0, "mm"),
          legend.position = "top",
          legend.direction = "horizontal")
  
  return(plot)
}
```

## Build plots based on P\<0.001 threshold

### CD4CD25

```{r plot CD4 CD25 inf}
signif <- c(17,2,7,15,20,14,9,5,6,16,4,12) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CD25 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD25")
  
CD4CD25
```

### CD4CCR7

```{r plot CD4 CCR7 inf}
signif <- c(21, 20, 18, 6, 7, 5, 19, 2, 9, 14, 17) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CCR7 <- plotinf(boxdata, jitterdata, aspect_ratio, "CCR7")
  
print(CD4CCR7)
```

### CD4CCR5

```{r plot CD4 CCR5 inf}
signif <- c(6, 14, 17, 10, 1, 13, 21, 9, 5) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CCR5 <- plotinf(boxdata, jitterdata, aspect_ratio, "CCR5")
  
print(CD4CCR5)
```

### CD4CD95

```{r plot CD4 CD95 inf}
signif <- c(17, 2, 14, 15) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CD95 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD95")
  
print(CD4CD95)
```

### CD4CD38

```{r plot CD4 CD38 inf}
signif <- c(14, 15, 10) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CD38 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD38")
  
print(CD4CD38)
```

### CD4CD28

```{r plot CD4 CD28 inf}
signif <- c(4, 19) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD4CD28 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD28")
  
print(CD4CD28)
```

### CD8CCR7

```{r plot CD8 CCR7 inf}
signif <- c(41, 39, 29, 40) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD8CCR7 <- plotinf(boxdata, jitterdata, aspect_ratio, "CCR7")
  
print(CD8CCR7)
```

### CD8CD25

```{r plot CD8 CD25 inf}
signif <- c(41, 39) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD8CD25 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD25")
  
print(CD8CD25)
```

### CD8CCR5

```{r plot CD8 CCR5 inf}
signif <- c(38, 30, 37, 41, 28, 40) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD8CCR5 <- plotinf(boxdata, jitterdata, aspect_ratio, "CCR5")
  
print(CD8CCR5)
```

### CD8CD28

```{r plot CD8 CD28 inf}
signif <- c(23, 22, 40) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD8CD28 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD28")
  
print(CD8CD28)
```

### CD8CD38

```{r plot CD8 CD38 inf}
signif <- c(32, 23, 22) # P<0.001 for either CT+ or CT+CoP at enrollment

aspect_ratio <- 3.6 / length(signif)
aspect_ratio

boxdata <- expressiondata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)   

jitterdata <- allupperqdata %>%
  filter(Infection == 0 | Infection == 1 | Infection == 2, 
         Visit == 0, 
         PopIndex %in% signif)

CD8CD38 <- plotinf(boxdata, jitterdata, aspect_ratio, "CD38")
  
print(CD8CD38)
```

# FU+ vs FU-

## Modify PopLabel to define line breaks for x-axis labels
```{r PopLabel line breaks}
PopLabelOrder <- PopLabelOrder %>% 
  str_replace_all("TEM ", "TEM\n") %>% 
  str_replace_all("TCM ", "TCM\n")

expressiondata <- expressiondata %>%
  mutate(PopLabel = str_replace_all(PopLabel, "TEM ", "TEM\n")) %>% 
  mutate(PopLabel = str_replace_all(PopLabel, "TCM ", "TCM\n")) %>% 
  mutate(PopLabel = factor(PopLabel, levels = PopLabelOrder)) 

allupperqdata <- allupperqdata %>%
  mutate(PopLabel = str_replace_all(PopLabel, "TEM ", "TEM\n")) %>% 
  mutate(PopLabel = str_replace_all(PopLabel, "TCM ", "TCM\n")) %>% 
  mutate(PopLabel = factor(PopLabel, levels = PopLabelOrder)) 
```

## Build plot function

```{r build function}
plotFU <- function(boxdata, jitterdata, y_variable) {
  plot <- boxdata %>%
    ggplot(aes(x = FollowUp, 
               y = !!sym(y_variable),
               fill = FollowUp)) +
    geom_boxplot(aes(alpha = Visit),
                 color = "black", 
                 outlier.shape = NA, 
                 coef = 0, 
                 size = 0.25) +
    scale_fill_manual(values = c("#00BE67", "#00A9FF"),
                      labels = c("FU-", "FU+")) +
    scale_alpha_manual(values = c(1, 0.4)) +
    geom_jitter(data = jitterdata, 
                aes(x = FollowUp, 
                    y = !!sym(y_variable),
                    color = interaction(FollowUp, Visit)),
                size = 0.1,
                position = position_jitterdodge(jitter.width = 1, 
                                                dodge.width = 0.75)) +  
    scale_color_manual(values = c("#006D3A", "#004972", "#008E4B", "#006799"), 
                       guide = "none") +
    coord_cartesian(ylim = c(NA, max(jitterdata[[y_variable]]))) + 
    facet_nested_wrap(~ PopLabel + Visit, 
                      nrow = 1, 
                      strip.position = "bottom",
                      nest_line = element_line(color = "black", linewidth = 0.5),
                      labeller = labeller(Visit = c("0" = "En", "1" = "1M"))) +
    guides(alpha = "none") +
    labs(y = y_variable) +
    theme_classic() +
    theme(aspect.ratio = 4,
          axis.title = element_text(size = 11, face = "bold"),  # Increase axis title font size
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.text = element_text(size = 11, color = "black"),
          strip.placement = "outside",
          strip.background = element_blank(),
          legend.text = element_text(size = 11),
          legend.title = element_blank(),
          legend.background = element_rect(fill = "transparent"),
          legend.box.margin = margin(0, 0, 0, 0),
          legend.spacing.x = unit(0, "mm"),
          legend.spacing.y = unit(0, "mm"),
          legend.position = "top",
          legend.direction = "horizontal")
  
  return(plot)
}
```

### Build plots based on P\<0.05 threshold (CT+ participants only)

## FUCD4CCR7
```{r plot FUCD4CCR7}
signif <- c(17, 19, 20, 12, 21, 16, 6) # P<0.05 for either CT+ or CT+CoP at enrollment

boxdata <- expressiondata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

jitterdata <- allupperqdata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

FUCD4CCR7 <- plotFU(boxdata, jitterdata, "CCR7")
  
print(FUCD4CCR7)
```

## FUCD8CCR7
```{r plot FUCD8CCR7}
signif <- c(31, 29, 32, 40) # P<0.05 for either CT+ or CT+CoP at enrollment

boxdata <- expressiondata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

jitterdata <- allupperqdata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

FUCD8CCR7 <- plotFU(boxdata, jitterdata, "CCR7")
  
print(FUCD8CCR7)
```

## FUCD4CCR6
```{r plot FUCD4CCR6}
signif <- c(9, 20, 17, 7) # P<0.05 for either CT+ or CT+CoP at enrollment

boxdata <- expressiondata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

jitterdata <- allupperqdata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

FUCD4CCR6 <- plotFU(boxdata, jitterdata, "CCR6")
  
print(FUCD4CCR6)
```

## FUCD8CD127
```{r plot FUCD8CD127}
signif <- c(22, 23) # P<0.05 for either CT+ or CT+CoP at enrollment

boxdata <- expressiondata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

jitterdata <- allupperqdata %>%
  filter(Infection == 1, 
         FollowUp != 2, 
         PopIndex %in% signif)  

FUCD8CD127 <- plotFU(boxdata, jitterdata, "CD127")
  
print(FUCD8CD127)
```
